Outline

### 2021-01-05 text/main.Rmd
more pruning

**Main Text:**
The flow of elements through biological communities fuels all life on Earth [@Welti2017]. Preserving these fluxes, often termed ecosystem functions, is critical for the integrity of ecosystems [@Welti2017].
For millennia, resources have been managed with an economic mindset to maximize desirable functions such as the production of biomass [@Weisser2017].
Sustaining multiple functions likely requires both high species richness and a variety of species assemblages across a landscape [@Zavaleta2010].
However, there can be trade-offs between functions, and efforts to maximize one function negatively impact another [@Zavaleta2010].
An understanding of such trade-offs is required to make informed management decisions, but simultaneously quantifying multiple ecosystem functions is challenging.
Therefore, trade-offs between functions, their drivers, and functional vulnerability are poorly understood in many ecosystems [@Brandl2019front].

Coral reefs are among the most diverse and productive ecosystems on Earth and provide essential ecosystem services to humanity [@Teh2013].
Yet, their integrity of is threatened by a plethora of anthropogenic stressors [@Hughes2017].
Over the past decades, severe declines in habitat quality and fish biomass have brought coral reef functioning to the forefront of scientific discourse [@Bellwood2019; @Brandl2019front].
However, our capacity to evaluate reef functioning primarily relies on static proxies of functions, such as live coral cover, standing stock biomass of reef fishes, or functional richness based on qualitative traits [e.g. @Cinner2020].
Conversely, we know comparatively little about elemental fluxes and their drivers (but see [@Allgeier2016]), which constitutes a severe limitation to management [@Brandl2019front].

@@

The five key ecosystem functions performed by fishes across the world's reefs exhibit high variability (Fig. 1).
Biomass is the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], and as can be expected from the cumulative nature of community-level functioning, we indeed observed a strong relationship between fish biomass and all five functions (Extended Data Fig. 1a-e, Extended Data Fig. 2).
However, our analyses demonstrate that reef fish functions vary remarkably  after accounting for biomass, as functions of communities with similar biomass may differ with two orders of magnitude and a two-fold difference in biomass can yield similar levels for functions (i.e. biomass production, Extended Data Fig. 1a-f).
Thus, using biomass as a proxy for functioning can mask differences in critical community-level functions.
Further, we demonstrate strong trade-offs among the five functions, independent of biomass (Fig.1, Extended Data Fig. 1g).
For example, high herbivory rates or nitrogen excretion negatively correlate with rates of phosphorus excretion.
Consequently, for a given value of standing stock biomass, no reef can yield above average values across all five functions.
While many reefs may stand out as hotspots for one function, none can maximize functioning as a holistic, multifaceted concept (Fig. 1).

Community structure and species-specific traits clearly impact rates of functioning.
First, using community-level ecological predictors known to affect elemental fluxes [@Schiettekatte2020] (body size, trophic level, species richness, biomass, temperature, and age structure; Fig. 2), we show that correlations between functions are mediated by contrasting aspects of community structure (Fig. 2; Extended Data Table 2; Extended Data Fig. 2).
For example, phosphorus excretion and piscivory are higher in communities with large-bodied fishes that occupy high trophic levels (Extended Fig. 1g; See @Allgeier2014).
In contrast, biomass production is highest in communities dominated by small and/or immature fishes at low trophic levels, creating a trade-off between biomass production and phosphorus excretion.
Metabolic theory predicts that small-bodied individuals have higher mass-specific metabolic rates, leading to elevated consumption rates and disproportional contributions to functions that rely on rapid energetic turnover [@Barneche2018; @Morais2020glob; @Brandl2019science].
Conversely, fishes in early life stages or with a nutrient-poor diet are often limited by phosphorus [@Schiettekatte2020], resulting in low contributions of these individuals to phosphorus excretion.
Thus, due to variations in organismal physiology and life-history traits [@Schiettekatte2020], fish community structure can significantly impact ecosystem-wide functioning [@Schramski2015].

Secondly, alongside features of community structure, ecosystem functioning may also be influenced by specific high-performing taxa.
Certain species or entire families tend to contribute more or less to a particular function, relative to their contribution to standing biomass (Fig. 3a; Extended Data Fig. 3,4).
Further, abundant species can impact rates of functioning at the community level [@Topor2019].
We show that functions consistently hinge on a few dominant species (Fig. 3b). Specifically, more than 50% of a given function, on average, is upheld by only 12% of the species present within a local community.
However, the identity of these species varies dramatically (Fig. 3c).
While few high-performing taxa dominate functioning in each location, there are no species that are important across their entire range, and most species are locally important in at least one community.
Indeed, 70% of all species contributed disproportionally to a specific function in at least one ommunity.
Despite high species richness on coral reefs, researchers often report the existence of functionally-dominant “key species” [e.g. @Bellwood2006].
Our results reveal that while functional dominance is indeed prevalent, the identity of local, dominant species vary strongly across different locations, suggesting that maintaining high levels of species richness across coral reef regions is essential to sustain global ecosystem functioning [@Topor2019].

@@

Similarly, the species-specific vulnerability of functionally-dominant species heavily affects functioning.
By combining species-level vulnerability scores to fishing and climate change induced coral loss [@Graham2011] and the contributions of each species to each function, we demonstrate that the loss of individuals most vulnerable to fishing will have greatest impacts on piscivory, followed by phosphorus excretion (Fig. 4).
Conversely, the loss of individuals due to climate change and consequent coral mortality may disproportionally reduce phosphorus excretion, nitrogen excretion, and biomass production.
Fishing and the loss of live coral impact species important for phosphorus excretion.
Surprisingly, although fishing pressure can negatively impact large herbivores such as parrotfishes [@Bellwood2011], herbivory is the least vulnerable function.
This may be due to the high variability in ecosystem roles within the comparatively large pool of herbivorous fish species.
While small herbivores are abundant and not particularly vulnerable to fishing, larger herbivorous species are frequently targeted and prone to functional extinction in areas with high fishing pressure [@Bellwood2012].
While herbivores of all body sizes and functional groups are combined in our assessment, their realized contributions to herbivory are strongly complementary, and it is important to recognize the specific functional roles performed by small subsets of herbivorous species (e.g. bioerosion [@Bellwood2011; @Green2009]).
Thus, our results reflect the overall vulnerability of functions; yet, this does not consider the distinct roles played by small groups of species nested within a broad function such as herbivory.

Conserving biomass, diversity, and ecosystem functioning are important objectives of most current conservation initiatives [@Cinner2020].
Even though safeguarding fish biomass enhances coral reef functioning, the strong trade-offs between key ecosystem functions reveals a critical challenge for coral reef conservation, where actions to enhance one function may negatively impact another.
For example, the establishment of marine protected areas, which are one of the primary conservation tactics for coral reefs [@Graham2020], may protect herbivorous species and thus provide benefits for herbivory. However, marine protected areas do not protect reefs from the pervasive effects of climate change [@Graham2020], and community shifts towards domination of herbivores may result in the decline of phosphorus excretion.
Thus, measuring conservation success with biomass or solely one function (e.g. herbivory) can mask the collapse of other essential functions.
It is necessary to gauge the state of reef ecosystems based on multiple, complementary, process-based functions [@Brandl2019front], as well as making informed decisions on local needs and stressors.
Finally, while there is a general consensus on the role of diversity in enhancing functioning [@Zavaleta2010], we highlight the importance of community structure and the identity of dominant species at the local scale.
While, maintaining the diversity of fishes is critical, at local scales, species richness only has a minor impact on individual functions. Rather, dissimilarity between local communities may be the most critical factor in maintaining functioning across the seascape since no species consistently provides high contributions for all functions or across its range [@Zavaleta2010].

>>>>

**Main Text:**
The flow of elements through biological communities fuels all life on Earth [@Welti2017]. Preserving these fluxes, often termed ecosystem functions, is critical for the integrity of ecosystems [@Welti2017].
For millennia, resources have been managed with an economic mindset to maximize desirable functions such as the production of biomass [@Weisser2017].
Sustaining multiple functions requires both high species richness and a variety of species assemblages across a landscape [@Zavaleta2010].
However, there can be trade-offs between functions, and efforts to maximize one function can negatively impact another [@Zavaleta2010].
An understanding of such trade-offs is required to make informed management decisions, but simultaneously quantifying multiple ecosystem functions is challenging.
Therefore, trade-offs between functions, their drivers, and functional vulnerability are poorly understood in many ecosystems [@Brandl2019front].

Coral reefs are among the most diverse and productive ecosystems on Earth and provide essential ecosystem services [@Teh2013].
Yet, their integrity of is threatened by a plethora of anthropogenic stressors [@Hughes2017].
Severe declines in habitat quality and fish biomass have brought coral reef functioning to the forefront of scientific discourse [@Bellwood2019; @Brandl2019front].
However, our capacity to evaluate reef functioning relies on static proxies of functions, such as live coral cover, standing stock biomass of reef fishes, or functional richness based on qualitative traits [e.g. @Cinner2020].
Conversely, we know comparatively little about elemental fluxes and their drivers (but see [@Allgeier2016]), which constitutes a severe limitation to management [@Brandl2019front].

@@

The five key ecosystem functions performed by fishes across the world's reefs exhibit high variability (Fig. 1).
Biomass is the most commonly employed indicator of coral reef functioning [@Cinner2020; @Brandl2019front], and we indeed observed a strong relationship between fish biomass and all five functions (Extended Data Fig. 1a-e, Extended Data Fig. 2).
However, our analyses demonstrate that functions vary remarkably after accounting for biomass - functions of communities with similar biomass may differ with two orders of magnitude or more (i.e. biomass production, Extended Data Fig. 1a-f).
Thus, using biomass as a proxy for functioning can mask differences in critical community-level functions.
Further, we demonstrate strong trade-offs among the five functions, independent of biomass (Fig.1, Extended Data Fig. 1g).
For example, high herbivory rates or nitrogen excretion negatively correlate with rates of phosphorus excretion.
Consequently, for a given value of biomass, no reef can yield above average values across all five functions.
While many reefs may stand out as hotspots for one function, none can maximize functioning as a holistic, multifaceted concept (Fig. 1).

Community structure and species-specific traits clearly impact rates of functioning.
First, using community-level ecological predictors known to affect elemental fluxes [@Schiettekatte2020] (body size, trophic level, species richness, biomass, temperature, and age structure; Fig. 2), we show that correlations between functions are mediated by contrasting aspects of community structure (Fig. 2; Extended Data Table 2; Extended Data Fig. 2).
For example, phosphorus excretion and piscivory are higher in communities with large-bodied fishes that occupy high trophic levels (Extended Fig. 1g; See @Allgeier2014).
In contrast, biomass production is highest in communities dominated by small and/or immature fishes at low trophic levels, creating a trade-off between biomass production and phosphorus excretion.
Metabolic theory predicts that small-bodied individuals have higher mass-specific metabolic rates, leading to elevated consumption rates and disproportional contributions to functions that rely on rapid energetic turnover [@Barneche2018; @Morais2020glob; @Brandl2019science].
Conversely, fishes in early life stages or with a nutrient-poor diet are often limited by phosphorus [@Schiettekatte2020], resulting in low contributions to phosphorus excretion.
Thus, due to variations in organismal physiology and life-history traits [@Schiettekatte2020], fish community structure can significantly impact ecosystem-wide functioning [@Schramski2015].

Secondly, alongside features of community structure, functions may also be influenced by specific high-performing taxa.
Certain species tend to contribute more or less to a particular function, relative to their contribution to biomass (Fig. 3a; Extended Data Fig. 3,4).
Further, abundant species impact rates of functioning at the community level [@Topor2019].
We show that functions consistently hinge on a few dominant species (Fig. 3b). Specifically, more than 50% of a given function, on average, is upheld by only 12% of the species present within a local community.
However, the identity of these species varies dramatically (Fig. 3c).
While few high-performing taxa dominate functioning in each location, there are no species that are important across their entire range, and most species are locally important in at least one community.
Indeed, 70% of all species contributed disproportionally to a specific function in at least one ommunity.
Despite high species richness on coral reefs, researchers often report the existence of functionally-dominant “key species” [e.g. @Bellwood2006].
Our results reveal that while functional dominance is indeed prevalent, the identity of local, dominant species vary strongly across different locations.

@@

Similarly, the species-specific vulnerability of functionally-dominant species heavily affects the vulnerability of functions.
By combining species-level vulnerability scores to fishing and climate change induced coral loss [@Graham2011] and the contributions of each species to each function, we illustrate that the loss of individuals most vulnerable to fishing will have greatest impacts on piscivory, followed by phosphorus excretion (Fig. 4).
Conversely, the loss of individuals due to climate change and consequent coral mortality may disproportionally reduce phosphorus excretion, nitrogen excretion, and biomass production.
Fishing and the loss of live coral impact species important for phosphorus excretion.
Surprisingly, although fishing pressure can negatively impact large herbivores such as parrotfishes [@Bellwood2011], herbivory is the least vulnerable function.
This may be due to the high variability in ecosystem roles within the comparatively large pool of herbivorous fish species.
While small herbivores are abundant and not particularly vulnerable to fishing, larger herbivorous species are frequently targeted and prone to functional extinction in areas with high fishing pressure [@Bellwood2012].
While herbivores of all body sizes and functional groups are combined in our assessment, their realized contributions to herbivory are strongly complementary, and it is important to recognize the specific functional roles performed by small subsets of herbivorous species (e.g. bioerosion [@Bellwood2011; @Green2009]).
Thus, our results reflect the overall vulnerability of functions; yet, this does not consider the distinct roles played by small groups of species nested within a broad function such as herbivory.

Conserving biomass, diversity, and ecosystem functions are important objectives of most current conservation initiatives [@Cinner2020].
Even though safeguarding fish biomass enhances functioning, the trade-offs between key functions reveal a critical challenge for coral reef conservation, where actions to enhance one function may negatively impact another.
For example, the establishment of marine protected areas, which are one of the primary conservation tactics for coral reefs [@Graham2020], may protect herbivorous species and thus provide benefits for herbivory. However, marine protected areas do not protect reefs from the pervasive effects of climate change [@Graham2020], and community shifts towards domination of herbivores may result in the decline of phosphorus excretion.
Thus, measuring conservation success with biomass or solely one function (e.g. herbivory) can mask the collapse of other essential functions.
It is necessary to gauge the state of reef ecosystems based on multiple, complementary, process-based functions [@Brandl2019front], as well as making informed decisions on local needs and stressors.
Finally, while there is a general consensus on the role of diversity in enhancing functioning [@Zavaleta2010], we highlight the importance of community structure and the identity of dominant species at the local scale.
While, maintaining the diversity of fishes is critical, at local scales, species richness only has a minor impact on individual functions. Rather, dissimilarity between local communities may be critical in maintaining functioning across the seascape since no species consistently provides high contributions for all functions or across its range [@Zavaleta2010].
